Batch plates with plyr
========================================================
Attempt to make the batch plates faster and better with plyr

First check and load libraries
```{r libraries}
libraries=c("nlstools","plyr","dplyr")
for(i in libraries){if(i %in% rownames(installed.packages()) == FALSE) {install.packages(i)}}
lapply(libraries, require, character.only=T)
```

define the equation we will fit
```{r init equation}

buchworm= od ~ od0 + (time >= lag) * (time <= (lag + (odmax - od0) *   
		log(10)/mumax)) * mumax * (time - lag)/log(10) + (time >= lag) *
		(time > (lag + (odmax - od0) * log(10)/mumax)) * (odmax - od0)
```

First function - read one directory of plates into a flat file

```{r dirfunction}
readonedir<-function(x){
  startingdir=getwd()
  setwd(x)
  filelist<-list.files()  											
	x3<-data.frame()															
	for(y in 1:length(filelist)){		
    time=strsplit(filelist[y],".c",fixed=T)[[1]][1]
		x1=data.frame(stringsAsFactors = FALSE)
    t=read.csv(filelist[y],header=TRUE,check.names=FALSE) 											
		t=t[1:13]
    for(zz in 2:length(t[1,])){
      for(xx in 1:length(t[,1])){
        x1=rbind(x1,data.frame(paste(t[xx,1],names(t)[zz],sep=""),t[xx,zz],time,x,stringsAsFactors = FALSE))
        }
      }
    x3=rbind(x3,as.data.frame(x1))
    }
  setwd(startingdir)
  names(x3)=c("well","od","time","plate")
  x3$time=as.numeric(x3$time)
  return(x3)
}
```

Now a couple of functions to fit the flat file
```{r fitfunction}
fitbuch<-function(x,y=1){
  x$od<-1-x$od
  if(y==1){x=x[x$time!=0,]}
  z=nls(buchworm,x,list(lag=35, mumax=0.025, od0 = 0.25, odmax = 0.95),control=nls.control(minFactor=1/4096,warnOnly=T))
  z3=data.frame()
  z3=rbind(z3,c(as.data.frame(t(coef(z))),sum(resid(z)^2),x$well[1],x$plate[1],as.numeric(strsplit(as.character(x$plate), split=' ', fixed=TRUE)[[1]][2])))
  names(z3)=c("lag","mumax","od0","odmax","residual","well","plate","run")
  return(z3)
}

plyrfit<-function(x){
  ddply(x,.(plate,well),function(df){fitbuch(df)})
}

dplyrfit<-function(x){z=x %.%
     group_by(plate, well) %.%
     do(fitbuch)
     return(do.call(rbind, z))
}

```

loop it over a directory - directory must have all names as "plate xx" and one file strainlist.csv

```{r looper}
looper<-function(x){
  startingdir=getwd()
  z3=data.frame()
  setwd(x)  			
  dirlist=list.files()
  for(zz in 1:(length(dirlist)-1)){
    z3=rbind(z3,plyrfit(readonedir(dirlist[zz])))
    }
  setwd(startingdir)
  return(z3)
}
```

next need to name them

```{r namer}
namer<-function(x,y){
  strainlist=read.csv(y)
  x$column=as.numeric(substring(as.character(x$well), 2, nchar(as.character(x$well))))
  x$row=substring(as.character(x$well), 1,1)
  z2=merge(x, strainlist, by=c("run","column"))
  asNumeric <- function(x){as.numeric(as.character(x))}
  return(z2)
}
```
put it all together!
```{r full thing}
setwd(choose.dir())
x=namer(looper(getwd()),"strainlist.csv")
```
next is to write it as a .csv - we can also remove the outliers and controls later
```{r output}
write.csv(x,file="outputfits.csv",row.names=F)
```
make a dataframe with all the means and sds
```{r compile all}
compileall<-function(x,y=0.1){
  x=x[which(x$residual<y),]
  z=ddply(x,.(strain,temperature),function(df){c(mean(df$lag),mean(df$mumax),mean(df$od0),mean(df$odmax),
                                             sd(df$lag),sd(df$mumax),sd(df$od0),sd(df$odmax),length(df$lag))})
  names(z)=c("strain","temperature","lag","mumax","od0","odmax","sdlag","sdmumax","sdod0","sdodmax","N")
  return(z)
}
```

```{r output}
y=compileall(x,0.1)
write.csv(y,file="grouped.csv",row.names=F)
```

Now some plots to check the residual cut offs
```{r plot by strain}
plotlots<-function(x,y,z){ #function of directory, strain, strainlist
  startdir=getwd()
  setwd(x)
	strainlist=read.csv(z)
	strainlist=strainlist[which(strainlist$strain==y),]
	for(i in 1:length(strainlist[,1])){
		singlehold=readonedir(paste("plate",sprintf("%02d", strainlist[i,]$run)))
		singlehold=singlehold[which(substring(singlehold$well, 2, nchar(x))==strainlist[i,]$column),]
		singlehold$temperature=strainlist[i,]$temperature
		singlehold  %.% group_by(well) %.% do(plotter)
	}
	setwd(startdir)
}


plotter<-function(x,well,strain){
  x=x[x$time!=0,]
	x$od=1-x$od
	workingfit<-nls(buchworm, x,list(lag=35, mumax=0.025, od0 = 0.25, odmax = 0.95),control=nls.control(minFactor=1/4096,warnOnly=T))
	plot(x$od~x$time,ylab="od",xlab="hours",main=c(paste("Plate:", x$plate[1],"Temp:", x$temperature[1],"Well",x$well[1],"Residual:",round(sum(resid(workingfit)^2),4))))
	points(x$time,predict(workingfit,list(time=x$time)),pch=3,col=2)
}
```
